---
title: "Input"
---

```{r include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
```

raylibr provides functions for keyboard, mouse, gamepad, and touch input. Input state is updated each frame, so these functions are meant to be called inside a game loop.

## Keyboard

Four functions check key state, each answering a different question:

| Function | True when... |
|----------|-------------|
| `is_key_pressed(key)` | Key went down **this frame** (fires once) |
| `is_key_down(key)` | Key is held down (fires every frame) |
| `is_key_released(key)` | Key went up **this frame** (fires once) |
| `is_key_up(key)` | Key is not held down (fires every frame) |

Use the `key$` enum for key codes:

```{r eval=FALSE}
library(raylibr)

run_game_loop(
  init_fn = function() {
    init_window(800L, 600L, "Keyboard Demo")
    set_target_fps(60L)
  },
  update_fn = function() {
    if (is_key_pressed(key$space)) message("Space pressed!")
    if (is_key_down(key$w)) message("W held down")

    begin_drawing()
    clear_background("black")
    draw_text("Press Space or hold W", 20L, 20L, 20L, "white")
    end_drawing()
  },
  cleanup_fn = close_window
)
```

For text input, use `get_key_pressed()` to consume queued key events or `get_char_pressed()` for Unicode characters:

```{r eval=FALSE}
char_code <- get_char_pressed()
while (char_code > 0L) {
  letter <- intToUtf8(char_code)
  char_code <- get_char_pressed()
}
```

By default, pressing Escape closes the window. Change or disable this with `set_exit_key()`:

```{r eval=FALSE}
set_exit_key(key$f12)  # close on F12 instead
set_exit_key(0L)       # disable exit key entirely
```

Common key codes: `key$space`, `key$enter`, `key$escape`, `key$up`, `key$down`, `key$left`, `key$right`, `key$a` through `key$z`.

## Mouse

```{r eval=FALSE}
pos <- get_mouse_position()  # Vector2: c(x, y)
x <- get_mouse_x()
y <- get_mouse_y()
delta <- get_mouse_delta()   # movement since last frame

if (is_mouse_button_pressed(mouse_button$left)) {
  message("Clicked at ", pos[1], ", ", pos[2])
}

wheel <- get_mouse_wheel_move()  # scroll amount (float)
```

Position and set the mouse cursor:

```{r eval=FALSE}
set_mouse_position(400L, 300L)
set_mouse_cursor(mouse_cursor$crosshair)
```

Available cursors: `mouse_cursor$default`, `mouse_cursor$arrow`, `mouse_cursor$ibeam`, `mouse_cursor$crosshair`, `mouse_cursor$pointing_hand`, `mouse_cursor$resize_ew`, `mouse_cursor$resize_ns`, `mouse_cursor$resize_all`, `mouse_cursor$not_allowed`.

## Cursor visibility

```{r eval=FALSE}
hide_cursor()
show_cursor()
disable_cursor()  # hides cursor and locks it for FPS-style controls
enable_cursor()
is_cursor_hidden()
is_cursor_on_screen()
```

## Gamepad

Check for connected gamepads and read their state:

```{r eval=FALSE}
if (is_gamepad_available(0L)) {
  name <- get_gamepad_name(0L)

  if (is_gamepad_button_down(0L, 7L)) {
    message("Right trigger pressed")
  }

  left_x <- get_gamepad_axis_movement(0L, 0L)
  left_y <- get_gamepad_axis_movement(0L, 1L)
}
```

The first argument is the gamepad index (0-based).

## Touch and gestures

For touch-enabled devices:

```{r eval=FALSE}
count <- get_touch_point_count()
pos <- get_touch_position(0L)  # first touch point

if (is_gesture_detected(4L)) {
  drag <- get_gesture_drag_vector()
}
```

## See also

For complete interactive input examples, see the [Mouse Test](demo-mouse.html) and [Snake](demo-snake.html) demos.
