---
title: "raymath"
---

```{r include=FALSE}
library(raylibr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
CAN_RENDER <- raylibr:::has_display()
```

raylibr includes the full raymath library — vector, matrix, and quaternion operations commonly needed in graphics programming. These use plain R types, so standard R operations work alongside raymath functions.

## R types as math types

| raymath type | R representation | Example |
|:-------------|:-----------------|:--------|
| Vector2 | `numeric(2)` | `c(1.0, 2.0)` |
| Vector3 | `numeric(3)` | `c(1.0, 2.0, 3.0)` |
| Vector4 / Quaternion | `numeric(4)` | `c(0, 0, 0, 1)` |
| Matrix | `matrix(nrow=4, ncol=4)` | `diag(4)` |

No wrapper classes — just plain R numerics and matrices.

## Float utilities

Basic math helpers that mirror GLSL built-in functions:

```{r float}
float_lerp(0, 10, 0.25)
float_clamp(15, 0, 10)
float_remap(5, 0, 10, 100, 200)
```

## Vector2 operations

Arithmetic, geometry, and interpolation on 2D vectors:

```{r vec2}
vector2_add(c(1, 2), c(3, 4))
vector2_length(c(3, 4))
vector2_normalize(c(3, 4))
vector2_distance(c(0, 0), c(3, 4))
vector2_dot_product(c(1, 0), c(0, 1))
vector2_lerp(c(0, 0), c(10, 10), 0.5)
vector2_rotate(c(1, 0), pi / 2)
```

Visualized — adding vectors `a` and `b`:

```{r vec2-visual, eval=CAN_RENDER}
raylibr_screenshot(function() {
  ox <- 200L; oy <- 150L; s <- 40

  draw_line(0L, oy, 400L, oy, "darkgray")
  draw_line(ox, 0L, ox, 300L, "darkgray")

  a <- c(3, 2)
  draw_line_ex(c(ox, oy), c(ox + a[1]*s, oy - a[2]*s), 3.0, "red")
  draw_text("a", as.integer(ox + a[1]*s + 5), as.integer(oy - a[2]*s), 20L, "red")

  b <- c(-1, 3)
  draw_line_ex(c(ox, oy), c(ox + b[1]*s, oy - b[2]*s), 3.0, "blue")
  draw_text("b", as.integer(ox + b[1]*s + 5), as.integer(oy - b[2]*s), 20L, "blue")

  ab <- vector2_add(a, b)
  draw_line_ex(c(ox, oy), c(ox + ab[1]*s, oy - ab[2]*s), 3.0, "green")
  draw_text("a+b", as.integer(ox + ab[1]*s + 5), as.integer(oy - ab[2]*s), 20L, "green")
})
```

Other useful Vector2 functions: `vector2_scale()`, `vector2_multiply()`, `vector2_negate()`, `vector2_reflect()`, `vector2_clamp()`, `vector2_move_towards()`.

## Vector3 operations

The same operations extend to 3D, plus a few new ones:

```{r vec3}
vector3_add(c(1, 2, 3), c(4, 5, 6))
vector3_cross_product(c(1, 0, 0), c(0, 1, 0))
vector3_length(c(1, 2, 2))
vector3_normalize(c(1, 2, 2))
```

`vector3_cross_product()` returns a vector perpendicular to both inputs — essential for computing surface normals and orientation.

```{r vec3-visual, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  draw_cube(c(0, 0.5, 0), 1, 1, 1, "red")

  pos <- vector3_add(c(0, 0.5, 0), c(2, 0, 0))
  draw_cube(pos, 1, 1, 1, "blue")

  draw_line_3d(c(0, 0.5, 0), pos, "yellow")
})
```

Other key functions: `vector3_rotate_by_axis_angle()`, `vector3_project()`, `vector3_reject()`, `vector3_unproject()`, `vector3_barycenter()`, `vector3_ortho_normalize()`.

## Matrix operations

Matrices represent transformations — translation, rotation, and scaling:

```{r matrix}
m <- matrix_identity()
m <- matrix_multiply(m, matrix_translate(1, 2, 3))
m <- matrix_multiply(m, matrix_rotate(c(0, 1, 0), pi / 4))
m <- matrix_multiply(m, matrix_scale(2, 2, 2))
```

Matrix multiplication chains transforms. The order matters: the rightmost transform is applied first.

Camera-related matrices:

```{r camera-matrix}
view <- matrix_look_at(c(4, 4, 4), c(0, 0, 0), c(0, 1, 0))
proj <- matrix_perspective(70 * pi / 180, 4/3, 0.1, 100)
```

Decompose a matrix back into its components:

```{r decompose}
m <- matrix_multiply(matrix_translate(1, 2, 3), matrix_scale(2, 2, 2))
parts <- matrix_decompose(m)
parts$translation
parts$scale
```

## Quaternions

Quaternions represent 3D rotations without gimbal lock. They're stored as `c(x, y, z, w)`:

```{r quat}
q1 <- quaternion_from_euler(0, pi/4, 0)
q2 <- quaternion_from_axis_angle(c(0, 1, 0), pi/2)

# Smooth interpolation between two rotations
q_mid <- quaternion_slerp(q1, q2, 0.5)

# Convert to a matrix for use with draw_model_ex() etc.
mat <- quaternion_to_matrix(q_mid)
```

`quaternion_slerp()` (Spherical Linear intERPolation) produces smooth rotation transitions — essential for animation.

Other quaternion functions: `quaternion_normalize()`, `quaternion_invert()`, `quaternion_multiply()`, `quaternion_from_matrix()`, `quaternion_to_euler()`, `quaternion_nlerp()`.
