---
title: "Shaders"
---

```{r include=FALSE}
library(raylibr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
CAN_RENDER <- raylibr:::has_display()
```

Shaders are small programs that run on the GPU and control how things are drawn. A **vertex shader** determines where vertices appear on screen, and a **fragment shader** determines the color of each pixel. raylibr lets you load custom GLSL shaders to create visual effects.

## Loading shaders

Load shaders from files or from strings in memory:

```r
# From files
shader <- load_shader("vertex.glsl", "fragment.glsl")

# Pass NULL for either to use Raylib's default
shader <- load_shader(NULL, "my_effect.glsl")

# From strings
fs_code <- "
#version 330
in vec2 fragTexCoord;
in vec4 fragColor;
out vec4 finalColor;
uniform sampler2D texture0;
void main() {
    vec4 color = texture(texture0, fragTexCoord) * fragColor;
    float gray = dot(color.rgb, vec3(0.299, 0.587, 0.114));
    finalColor = vec4(gray, gray, gray, color.a);
}
"
shader <- load_shader_from_memory(NULL, fs_code)
```

## Using shaders

Wrap your drawing calls in `begin_shader_mode()` / `end_shader_mode()`:

```r
begin_shader_mode(shader)
draw_texture(my_texture, 0L, 0L, "white")
end_shader_mode()
```

Everything drawn between these calls uses the custom shader instead of the default rendering.

## Uniforms

Pass data from R to the shader via **uniforms** — named variables in the GLSL code:

```r
loc <- get_shader_location(shader, "intensity")
set_shader_value(shader, loc, 0.5)           # float
set_shader_value(shader, loc, c(1.0, 0.5))   # vec2
set_shader_value(shader, loc, c(1, 0, 0))    # vec3
set_shader_value(shader, loc, c(1, 0, 0, 1)) # vec4
```

`set_shader_value()` automatically picks the right GLSL type based on the length of the R vector.

For matrices, use `set_shader_value_matrix()`. For textures, use `set_shader_value_texture()`.

## Post-processing pattern

The most common shader use case is post-processing: render the scene normally to an off-screen texture, then draw that texture to the screen through a shader.

```r
# Setup
target <- load_render_texture(400L, 300L)
shader <- load_shader(NULL, "my_effect.glsl")
source_rec <- rectangle(0, 0, 400, -300)  # flipped Y for render texture

# In the game loop:
begin_texture_mode(target)
  clear_background("black")
  # ... draw your scene normally ...
end_texture_mode()

begin_drawing()
  begin_shader_mode(shader)
  draw_texture_rec(target$texture, source_rec, c(0, 0), "white")
  end_shader_mode()
end_drawing()
```

Note the negative height in the source rectangle — render textures are vertically flipped in OpenGL.

## Example

Without a shader, a scene with colored circles looks like this:

```{r no-shader, eval=CAN_RENDER}
raylibr_screenshot(function() {
  draw_circle(100L, 150L, 80.0, "red")
  draw_circle(200L, 150L, 80.0, "green")
  draw_circle(300L, 150L, 80.0, "blue")
  draw_text("No shader", 10L, 10L, 20L, "white")
})
```

A grayscale fragment shader would convert this to shades of gray by computing a weighted luminance from the RGB channels. The GLSL code uses the standard luminance weights `(0.299, 0.587, 0.114)` to match human perception.

## Further reading

See these demos for more advanced shader usage:

- [Shader Stars](demo-shader.html) — animated GLSL fragment shader with uniform updates each frame
- [Julia Sets](demo-julia.html) — fractal rendering with mouse-driven zoom via shader uniforms
