---
title: "Getting Started"
---

```{r include=FALSE}
library(raylibr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
CAN_RENDER <- raylibr:::has_display()
```

## Installation

raylibr is not yet on CRAN. Install the development version from GitHub:

```r
pak::pak("jeroenjanssens/raylibr")
```

raylibr works on macOS, Linux, and Windows. It also runs in the browser via [webR](guide-webr.html).

## Your first image

The quickest way to see raylibr in action is `raylibr_screenshot()`. It creates a hidden window, runs your drawing code, and returns the result as an image:

```{r first-image, eval=CAN_RENDER}
raylibr_screenshot(function() {
  draw_circle(200L, 150L, 80.0, "yellow")
})
```

The default canvas is 400x300 pixels with a black background. Everything you draw appears on top.

You can compose multiple shapes in a single screenshot:

```{r composition, eval=CAN_RENDER}
raylibr_screenshot(function() {
  draw_rectangle(50L, 50L, 300L, 200L, "midnightblue")
  draw_circle(200L, 150L, 60.0, "gold")
  draw_triangle(
    c(200, 70), c(140, 190), c(260, 190),
    color_alpha("white", 0.5)
  )
  draw_text("raylibr", 145L, 220L, 30L, "white")
})
```

## The game loop

Raylib renders frame by frame: each iteration of a loop draws one frame to the screen. This is different from R's typical batch execution model.

The classic pattern looks like this:

```r
init_window(800L, 600L, "My Game")
set_target_fps(60L)

while (!window_should_close()) {
  begin_drawing()
  clear_background("black")
  draw_text("Hello!", 300L, 280L, 40L, "white")
  end_drawing()
}

close_window()
```

For code that needs to work both on desktop and in the browser (via webR), use `run_game_loop()` instead of a `while` loop:

```r
run_game_loop(
  init_fn = function() {
    init_window(800L, 600L, "My Game")
    set_target_fps(60L)
  },
  update_fn = function() {
    begin_drawing()
    clear_background("black")
    draw_text("Hello!", 300L, 280L, 40L, "white")
    end_drawing()
  },
  cleanup_fn = close_window
)
```

## Conventions

raylibr follows a few consistent conventions:

**snake_case names.** Raylib's `DrawCircle()` becomes `draw_circle()`.

**Integers need `L`.** Width, height, pixel coordinates, and font sizes are integers. Write `200L`, not `200`.

**R color names work everywhere.** Any function that takes a color accepts R's 657 built-in color names: `"red"`, `"steelblue"`, `"grey90"`, etc. See the [Colors](guide-colors.html) guide for details.

**Vectors map to R numerics.** A Raylib `Vector2` is `c(x, y)`, a `Vector3` is `c(x, y, z)`, and a `Vector4` is `c(x, y, z, w)`. A Raylib `Matrix` is a 4x4 R matrix.

## Structs

Some Raylib types have dedicated constructors:
```{r structs}
rec <- rectangle(10, 20, 300, 200)
rec$width

cam <- camera_3d(c(4, 4, 4))
cam$fovy

bb <- bounding_box(c(-1, -1, -1), c(1, 1, 1))
bb$min
```

Other constructors include `color()`, `ray()`, `camera_2d()`, `ray_collision()`, and `transform()`.

Check types with predicates:
```{r type-check}
is_rectangle(rec)
is_color("red")
```

## Enums

Raylib enums are available as named lists:

```{r enums}
flag$window_resizable
key$space
mouse_button$left
camera_mode$orbital
```

Use them with functions like `set_config_flags(flag$window_resizable)` or `is_key_pressed(key$space)`.

## What's next

- [Drawing 2D Shapes](guide-drawing-2d.html) --- circles, rectangles, lines, and more
- [Drawing 3D](guide-drawing-3d.html) --- cubes, spheres, and cameras
- [Colors](guide-colors.html) --- R color names, transparency, HSV, and blending
- [Text & Fonts](guide-text.html) --- rendering text and loading custom fonts
- [Images & Textures](guide-images.html) --- loading, generating, and drawing images
- [Input](guide-input.html) --- keyboard, mouse, and gamepad
- [Audio](guide-audio.html) --- sound effects and music
- [Camera](guide-camera.html) --- 2D and 3D camera systems
- [Shaders](guide-shaders.html) --- GLSL shaders and post-processing
- [raygui](guide-raygui.html) --- immediate-mode GUI widgets
- [raymath](guide-raymath.html) --- vector, matrix, and quaternion math
- [webR](guide-webr.html) --- running raylibr in the browser
