---
title: "Drawing 3D"
---

```{r include=FALSE}
library(raylibr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
CAN_RENDER <- raylibr:::has_display()
```

This guide covers 3D shape rendering. Every example uses `raylibr_screenshot_3d()`, which sets up a hidden window with a camera at `c(4, 4, 4)` looking at the origin, draws a reference grid, then runs your drawing code.

## 3D coordinate system

Raylib uses a **right-handed** coordinate system: X points right, **Y points up**, and Z points toward you. This is different from the 2D system where Y points down.

## Camera

A 3D scene needs a camera. `camera_3d()` creates one:

```{r camera-struct}
cam <- camera_3d(
  position = c(4, 4, 4),
  target = c(0, 0, 0),
  up = c(0, 1, 0),
  fovy = 70.0,
  projection = camera_projection$perspective
)
cam$position
```

The defaults (`target = c(0,0,0)`, `up = c(0,1,0)`, `fovy = 70`) work well for most cases, so `camera_3d(c(4, 4, 4))` is usually enough. See the [Camera](guide-camera.html) guide for modes, controls, and coordinate conversion.

## The 3D drawing block

3D drawing calls must be wrapped in `begin_mode_3d()` / `end_mode_3d()`:

```r
begin_drawing()
clear_background("black")
begin_mode_3d(cam)
draw_cube(c(0, 0, 0), 2.0, 2.0, 2.0, "red")
end_mode_3d()
end_drawing()
```

`raylibr_screenshot_3d()` handles all this boilerplate --- you just provide the drawing code:

```{r simple-3d, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  draw_cube(c(0, 0.5, 0), 1.0, 1.0, 1.0, "tomato")
})
```

## Cubes

```{r cubes, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  draw_cube(c(-1.5, 0.5, 0), 1.0, 1.0, 1.0, "steelblue")
  draw_cube_wires(c(-1.5, 0.5, 0), 1.0, 1.0, 1.0, "white")
  draw_cube(c(1.5, 0.75, 0), 1.5, 1.5, 1.5, color_alpha("gold", 0.7))
  draw_cube_wires(c(1.5, 0.75, 0), 1.5, 1.5, 1.5, "white")
})
```

`draw_cube()` takes a center position and width/height/length as floats. `draw_cube_wires()` draws the wireframe outline. `draw_cube_v()` and `draw_cube_wires_v()` accept size as a `Vector3`.

## Spheres

```{r spheres, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  draw_sphere(c(-1.5, 1, 0), 1.0, "tomato")
  draw_sphere_wires(c(1.5, 1, 0), 1.0, 12L, 12L, "dodgerblue")
  draw_sphere_ex(c(0, 1, -2), 0.7, 16L, 16L, "mediumseagreen")
})
```

`draw_sphere()` takes a center and radius. `draw_sphere_ex()` adds control over the number of rings and slices (more = smoother). `draw_sphere_wires()` draws a wireframe sphere.

## Cylinders and capsules

```{r cylinders, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  draw_cylinder(c(-2, 0, 0), 0.5, 0.5, 2.0, 16L, "steelblue")
  draw_cylinder_ex(c(0, 0, 0), c(0, 2, 0), 0.8, 0.3, 16L, "gold")
  draw_capsule(c(2, 0, 0), c(2, 2, 0), 0.5, 8L, 8L, "tomato")
})
```

`draw_cylinder()` takes a base position, top/bottom radii, height, and slices. `draw_cylinder_ex()` takes start and end points for arbitrary orientation. `draw_capsule()` is a cylinder with rounded ends.

## Planes and lines

```{r primitives, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  draw_plane(c(0, 0, 0), c(3, 3), color_alpha("steelblue", 0.5))
  draw_line_3d(c(-2, 0, -2), c(2, 3, 2), "tomato")
  draw_point_3d(c(0, 2, 0), "gold")
  draw_triangle_3d(
    c(-1, 0, 1), c(1, 0, 1), c(0, 2, 1),
    "mediumseagreen"
  )
})
```

`draw_plane()` draws a flat quad at a given position with a given size.

## Grid

`draw_grid()` renders a reference grid on the XZ plane, centered at the origin. `raylibr_screenshot_3d()` includes `draw_grid(10L, 1.0)` automatically:

```{r grid, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  # The grid is already drawn by raylibr_screenshot_3d()
  # Just add something to see the scale
  draw_cube(c(0, 0.5, 0), 1.0, 1.0, 1.0, color_alpha("tomato", 0.6))
  draw_cube(c(2, 0.5, 0), 1.0, 1.0, 1.0, color_alpha("steelblue", 0.6))
})
```

Each grid square is 1 unit wide. The cubes above are 1 unit on each side, spaced 2 units apart.

## Bounding boxes

`bounding_box()` defines an axis-aligned box by its min and max corners. Useful for collision detection and visualization:

```{r bounding-box, eval=CAN_RENDER}
raylibr_screenshot_3d(function() {
  bb <- bounding_box(c(-1, 0, -1), c(1, 2, 1))
  draw_bounding_box(bb, "tomato")
  draw_sphere(c(0, 1, 0), 0.5, color_alpha("gold", 0.5))
})
```

## Models

For complex 3D objects, load models from files with `load_model()` and draw them with `draw_model()`. See the [3D Model demo](demo-model.html) for a complete example with textures and post-processing shaders.
