---
title: "raygui"
---

```{r include=FALSE}
library(raylibr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
CAN_RENDER <- raylibr:::has_display()
```

raygui is an **immediate-mode** GUI library bundled with raylibr. Unlike retained-mode GUI toolkits (where you create widgets once and query their state), immediate-mode GUIs call a function for each control every frame. The function draws the control and returns its current value or interaction state.

## Layout

Every control is positioned with a `rectangle(x, y, width, height)`. There is no automatic layout — you place each control explicitly.

```{r layout, eval=CAN_RENDER}
raylibr_screenshot(function() {
  gui_panel(rectangle(20, 20, 360, 260), "Settings")
  gui_label(rectangle(40, 60, 100, 30), "Name:")
  gui_text_box(rectangle(140, 60, 200, 30), "raylibr", 64L, FALSE)
  gui_label(rectangle(40, 100, 100, 30), "Volume:")
  gui_slider_bar(rectangle(140, 100, 200, 30), "", "80%", 0.8, 0, 1)
  gui_check_box(rectangle(40, 140, 20, 20), "Enable sound", TRUE)
  gui_button(rectangle(140, 200, 100, 40), "Apply")
})
```

## Buttons and labels

`gui_button()` draws a clickable button and returns `TRUE` when clicked:

```r
if (gui_button(rectangle(100, 100, 120, 40), "Click me")) {
  # handle click
}
```

`gui_label()` draws static text. `gui_label_button()` is a clickable text label.

## Toggles and checkboxes

```{r toggles, eval=CAN_RENDER}
raylibr_screenshot(function() {
  gui_toggle(rectangle(40, 40, 120, 30), "Option A", TRUE)
  gui_toggle(rectangle(40, 80, 120, 30), "Option B", FALSE)
  gui_toggle_group(rectangle(40, 130, 100, 30), "Low;Medium;High", 1L)
  gui_check_box(rectangle(40, 180, 20, 20), "Fullscreen", FALSE)
  gui_check_box(rectangle(40, 210, 20, 20), "VSync", TRUE)
})
```

`gui_toggle()` returns the new active state. `gui_toggle_group()` shows multiple options and returns the selected index. `gui_check_box()` returns the new checked state.

## Sliders and progress bars

```{r sliders, eval=CAN_RENDER}
raylibr_screenshot(function() {
  gui_slider(rectangle(40, 40, 300, 20), "0", "100", 65, 0, 100)
  gui_slider_bar(rectangle(40, 80, 300, 20), "Vol", "", 0.4, 0, 1)
  gui_progress_bar(rectangle(40, 120, 300, 20), "", "75%", 0.75, 0, 1)
  gui_spinner(rectangle(40, 160, 150, 30), "", 42L, 0L, 100L, FALSE)
})
```

`gui_slider()` and `gui_slider_bar()` return the new value. `gui_progress_bar()` is display-only. `gui_spinner()` shows a value with +/- buttons.

## Text input

`gui_text_box()` renders an editable text field. The `edit_mode` parameter controls whether it's active for input:

```r
text <- gui_text_box(rectangle(40, 40, 200, 30), text, 256L, edit_mode)
```

`gui_combo_box()` shows a dropdown-style selector, `gui_dropdown_box()` shows an expandable list.

## Color controls

```{r color, eval=CAN_RENDER}
raylibr_screenshot(function() {
  gui_color_picker(rectangle(40, 20, 150, 150), "", color(200L, 100L, 50L))
  gui_color_bar_hue(rectangle(210, 20, 30, 150), "", 120)
  gui_color_bar_alpha(rectangle(260, 20, 30, 150), "", 0.7)
})
```

`gui_color_picker()` returns a color. `gui_color_bar_hue()` returns a hue value (0-360). `gui_color_bar_alpha()` returns an alpha value (0-1).

## Containers

- `gui_panel(bounds, text)` — a bordered panel with an optional title
- `gui_group_box(bounds, text)` — a labeled group outline
- `gui_window_box(bounds, text)` — a window with a close button (returns `TRUE` when closed)
- `gui_scroll_panel(bounds, text, content, scroll, view)` — scrollable area for content larger than the visible region

## Styling

Customize the look of controls with `gui_set_style()`:

```r
# Set a property for a specific control type
gui_set_style(gui_control$button, gui_control_property$text_color_normal, 0xFF0000FF)

# Load a complete style file
gui_load_style("my_style.rgs")

# Reset to defaults
gui_load_style_default()
```

State functions control whether controls are interactive:

```r
gui_disable()  # gray out all controls
gui_enable()   # re-enable
gui_lock()     # controls visible but non-interactive
gui_unlock()
```

## Further reading

See the [GUI Controls demo](demo-gui.html) for a complete interactive example with all major control types.
