---
title: "Text & Fonts"
---

```{r include=FALSE}
library(raylibr)
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
CAN_RENDER <- raylibr:::has_display()
```

raylibr provides several functions for drawing text, from simple one-liners to fully controlled rendering with custom fonts, rotation, and spacing.

## Simple text

`draw_text()` is the quickest way to put text on screen. Pass the string, position, font size (integer), and color.

```{r text-sizes, eval=CAN_RENDER}
raylibr_screenshot(function() {
  draw_text("Small (10)", 20L, 20L, 10L, "lightgray")
  draw_text("Medium (20)", 20L, 50L, 20L, "skyblue")
  draw_text("Large (40)", 20L, 90L, 40L, "yellow")
  draw_text("Huge (60)", 20L, 150L, 60L, "tomato")
})
```

The default font is a built-in bitmap font included with Raylib.

## Custom fonts

For higher quality text, load a TTF or OTF font with `load_font_ex()`:

```{r eval=FALSE}
font <- load_font_ex("path/to/font.ttf", 32L)
draw_text_ex(font, "Custom font!", c(100, 100), 32.0, 2.0, "white")
unload_font(font)
```

`draw_text_ex()` takes a font object, text, position (Vector2), font size (float), character spacing, and color.

## Rotated text

`draw_text_pro()` adds rotation and an origin point. The origin is relative to the top-left of the text and determines the pivot for rotation.

```{r text-rotation, eval=CAN_RENDER}
raylibr_screenshot(function() {
  font <- get_font_default()
  draw_text_pro(font, "0 degrees", c(200, 150), c(0, 0), 0.0, 20.0, 2.0, as_color("white"))
  draw_text_pro(font, "45 degrees", c(200, 150), c(0, 0), 45.0, 20.0, 2.0, as_color("lime"))
  draw_text_pro(font, "90 degrees", c(200, 150), c(0, 0), 90.0, 20.0, 2.0, as_color("skyblue"))
  draw_text_pro(font, "-30 degrees", c(200, 150), c(0, 0), -30.0, 20.0, 2.0, as_color("yellow"))
  draw_circle(200L, 150L, 4.0, "red")
})
```

The red dot marks the shared origin point. Each string fans out from there at a different angle.

## Measuring text

`measure_text()` returns the width of a string in pixels at a given font size. Use it to center text:

```{r text-center, eval=CAN_RENDER}
raylibr_screenshot(function() {
  text <- "Centered!"
  size <- 40L
  w <- measure_text(text, size)
  x <- (400L - w) %/% 2L
  draw_text(text, x, 130L, size, "gold")
  draw_line(200L, 0L, 200L, 300L, "darkgray")
})
```

For custom fonts, use `measure_text_ex(font, text, font_size, spacing)` which returns a Vector2 with both width and height.

## Line spacing

`set_text_line_spacing()` controls the vertical distance between lines when drawing multi-line strings (strings containing `\n`):

```{r eval=FALSE}
set_text_line_spacing(24L)
draw_text("Line one\nLine two\nLine three", 20L, 20L, 20L, "white")
```

## Text utilities

raylibr includes string manipulation functions:

```{r eval=FALSE}
text_to_upper("hello")       # "HELLO"
text_to_lower("HELLO")       # "hello"
text_to_snake("HelloWorld")  # "hello_world"
text_to_pascal("hello_world") # "HelloWorld"
text_to_camel("hello_world") # "helloWorld"
text_subtext("Hello", 0L, 3L) # "Hel"
text_length("Hello")          # 5
```
